---
title: "Practical Machine Learning"
author: "Javier González Onrubia"
date: "28 de marzo de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment. 

Your submission for the Peer Review portion should consist of a link to a Github repo with your R markdown and compiled HTML file describing your analysis. Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. It will make it easier for the graders if you submit a repo with a gh-pages branch so the HTML page can be viewed online (and you always want to make it easy on graders :-).

- Métodos de clasificación
- Rattle -> Decission Trees, Random Forest
- Matriz de confusión

## Notes

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment. 

```{r finalcode}
library(caret)
library(parallel)
library(doParallel)

set.seed(12345)

## Loading files
download.file(
    "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",
    "pml-training.csv", method = "curl")
download.file(
    "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",
    "pml-testing.csv", method = "curl")

training <- read.csv("pml-training.csv", 
                     na.strings  = c("", "NA", "#DIV/0!"))
testing  <- read.csv("pml-testing.csv")

## Data preprocesing and cleaning
### At least 50% of non-NA data available
training <- training[,colSums(!is.na(training))>=(nrow(training)*0.5)]
### Pruning columns considered not useful as predictor
### (X, username, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, new_window, num_window)
training <-training[, -c(1:7)]

## Subsetting training data into training and testing sets
inTraining <- createDataPartition(training$classe, p = 0.75, list = FALSE)
trainset <- training[inTraining,]
testset <- training[-inTraining,]

## Parallelizing
cluster <- makeCluster(detectCores()-1)
registerDoParallel(cluster)

## Classifying
fitControl <- trainControl(method = "cv", 
                           number = 5, 
                           allowParallel = TRUE)

m.rf <- train(classe~., 
              method = "rf", 
              data = trainset, 
              trControl = fitControl)

m.rf.pca    <- train(classe~., data = trainset, method = "rf", preProcess = c("pca"))
m.lda       <- train(classe~., data = trainset, method = "lda")
m.lssvmPoly <- train(classe~., data = trainset, method = "lssvmPoly")
m.nb        <- train(classe~., data = trainset, method = "nb")
m.gamboost  <- train(classe~., data = trainset, method = "gamboost")

stopCluster(cluster)
registerDoSEQ()

## Prediction on testing
p.rf <- predict(m.rf, testset)
p.rf.pca <- predict(m.rf.pca, testset)
p.lda <- predict(m.lda, testset)
p.lssvmPoly <- predict(m.lssvmPoly, testset)
p.nb <- predict(m.nb, testset)
p.gamboost <- predict(m.gamboost, testset)
## Cross-validation
confusionMatrix(p.rf, testset$classe)

## Classe predictions for testing data
p.rf(m.rf, testing)

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
