---
title: "Practical Machine Learning"
author: "Javier González Onrubia"
date: "28 de marzo de 2018"
output: html_document
---

```{r setup, include=FALSE, results = "hide"}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
Sys.setlocale(locale = "en_US.UTF-8")
```

## Summary
This project uses data from a Human Activity Recognition (HAR) project that tries 
to classify how well does an individual do a given exercise
(http://groupware.les.inf.puc-rio.br/har). Our goal is find a classifier model
that predicts how well (the `classe` variable) the exercise was done.

This project uses 2 different models based on classification trees and random
forests, finding out the latter to be a best fit for prediction.

Additionally, a brief analysis on the random forest model is done, showing which predictors explain the most of the variability and how many of them give the better results.

And finally, a 20-case testing data is used to predict the `classe` that each
observation belongs to.

## Data Cleaning 

First of all, we load the training data and the testing one, the latter being
used as the quiz testing set.

```{r dataloading, cache = TRUE}
download.file(
    "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",
    "pml-training.csv", method = "curl")
download.file(
    "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",
    "pml-testing.csv", method = "curl")

training <- read.csv("pml-training.csv", 
                     na.strings  = c("", "NA", "#DIV/0!"))
quiz  <- read.csv("pml-testing.csv")
```

We clean the columns with a very high level of NA values (>=50%) as they
won't be useful as predictors. Other descriptive columns as X, username, 
measurement date/time and informational ones as new_window and num_window are
also removed as predictors.

```{r datacleaning, cache = TRUE}
# At least 50% of non-NA data available
training <- training[,colSums(!is.na(training))>=(nrow(training)*0.5)]
# Pruning columns considered not useful as predictors
training <-training[, -c(1:7)]
```

Analyzing which candidate predictors could have near-zero variance, we find no
predictor that could be set apart using this criteria.

```{r nzv}
library(caret)
nearZeroVar(training, saveMetrics = TRUE)
```

Finally, we'll be using `r ncol(training) - 1` predictors.

## Subsetting data into training and test set

As the `quiz` data can't be used as a testing set, we split 25% os the training
data to be used as the testing set. We call this set as the `holdout` set since
as we'll be using cross-validation, each training sample will be tested against
the test sample of each of the folds.

```{r setpartition, cache = TRUE}
set.seed(12345)

# Subsetting training data into training and testing sets
inTraining <- createDataPartition(training$classe, p = 0.75, list = FALSE)
trainset <- training[inTraining,]
holdout <- training[-inTraining,]
```

This subsetting gives as for `r nrow(trainset)` observations for the training 
set and `r nrow(holdout)` observations for the holdout one.

### Parallelizing

To increase training performance the `parallel` and `doParallel` packages were
used as follows:

```{r parallelsetup}
library(parallel)
library(doParallel)
cluster <- makeCluster(detectCores()-1)
registerDoParallel(cluster)
```

## Classifying

As for classifying methods, we'll be using classification trees and random 
forests, both of them using 5-fold cross-validation. To build the model we'll
use the `trainset` data and the `holdout` as the testing set to validate
how good they fit the data not considered in building the model.

```{r classifying, cache = TRUE}
fitControl <- trainControl(method = "cv", number = 5, allowParallel = TRUE)

m.rf <- train(classe~., method = "rf", data = trainset, trControl = fitControl)
m.rpart <- train(classe~., method = "rpart", data = trainset, trControl = fitControl)
stopCluster(cluster)
registerDoSEQ()
```

Looking at the accuracy for the 5-fold cross-validation:

```{r accuracy}
m.rpart$results
m.rf$results
```

We get a `r round(sort(m.rpart$results$Accuracy, decreasing = TRUE)[1]*100, digits = 2)`% accuracy
for the classification tree and a 
`r round(sort(m.rf$results$Accuracy, decreasing = TRUE)[1]*100, digits = 2)`%
accuracy for the random forest method. As the latter gives us the best accuracy,
we'll select this method for further predictions.

Nevertheless, the confusion matrix for both methods using the `holdout` data
as a best-fit evaluator confirms the random forest as the best model.

```{r confmatrix.rpart}
confusionMatrix(predict(m.rpart, holdout), holdout$classe)
plot(predict(m.rpart, holdout), holdout$classe)
```

```{r confmatrix.rf}
confusionMatrix(predict(m.rf, holdout), holdout$classe)
plot(predict(m.rf, holdout), holdout$classe)
```

## The Random forest model

```{r rfmodel}
m.rf
```
The model summary shows that the best accuracy is about 
`r round(sort(m.rf$results$Accuracy, decreasing = TRUE)[1]*100, digits = 2)`% 
using `r m.rf$bestTune` variables (`mtry`), so the OOB error would be about
`r 1 - round(sort(m.rf$results$Accuracy, decreasing = TRUE)[1]*100, digits = 2)`%

```{r rsp}
plot(m.rf)
```

As for the predictor importance:

```{r predictorimportance}
library(randomForest)
library(dplyr)
pred_importance <- data.frame(
                        predictor = setdiff(colnames(trainset), "classe"),
                        importance=as.vector(importance(m.rf$finalModel))
        )
pred_importance <- arrange(pred_importance, importance)
pred_importance$predictor <- factor(pred_importance$predictor,
                                   levels=pred_importance$predictor)

p <- ggplot(pred_importance, aes(x=predictor, weight=importance, fill=predictor))
p <- p + geom_bar() + coord_flip()
p <- p + ggtitle("Predictor Importance from Random Forest Fit")
p <- p + xlab("Predictor") + ylab("Importance (Mean Decrease in Gini Index)")
p <- p + scale_fill_discrete(name="Predictor")
p + theme(axis.text.x=element_blank(),
          axis.text.y=element_text(size=8),
          axis.title=element_text(size=14),
          plot.title=element_text(size=16),
          legend.position = "none")
```

## Quiz prediction

```{r quizprediction, cache = TRUE}
predict(m.rf, quiz)
```
